Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-07-22 GWLDCI.tex
paper 1 readthrough

In \cite{wilsonloop} the first author and Amat showed that every admissible Wilson loop diagrams can be identified with a unique positroid cell, and observed that this mapping is neither one-to-one nor onto. In this paper we give a simple characterization of this lack of injectivity in terms of an equivalence relation on the Wilson loop diagrams, and enumerate the number of different Wilson loop diagrams that correspond to each positroid cell. We also study the lack of surjectivity, and show that the number of distinct positroid cells associated to Wilson loop diagrams of a given size is given by counting the directions in an associahedron.

The second paper \cite{generalcombinatoricsII} in this series identifies precisely which positroid cell is associated to a given Wilson loop diagram, by giving an algorithm for passing from a Wilson loop diagram to the Grassmann necklace of the associated positroid cell. In this manner we give a direct path from Wilson loop diagram to positroid cell circumventing the painstaking process of examining bases sets. With this identification in hand, we also prove that any Wilson loop diagram with $k$ propagators corresponds to a $3k$ dimensional positroid cell, and characterize the volume forms that Wilson loop diagrams associate to each positroid cell.

@@

Meanwhile, there is a separate body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. As noted above, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of the positive Grassmannian $\Gr(k,n)$ \cite{wilsonloop}. In (separate) papers with the second author and Marcott, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmannians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

However, the two approaches are not identical: for example, in the case of Wilson loop diagrams with $k=2$ and $n=6$, past work of some of us shows that the manifold is not contractible \cite{casestudy}. More generally, it has been shown that the spaces defined by the Amplituhedron and by Wilson loop diagrams are not the same \cite{Amplituhedronsquared}. In fact, it is in some ways more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams; in this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}. \note{Check with Cam.}

@@

Wherever possible, we drop the subscrips on propagators and their identifying vertices. Propagators are undirected, so $p = (i,j) = (j,i)$. In previous work, we have used the convention, we write $p = (i,j)$ with  $i +1 \leq j$ relative to the first vertex.

We depict this data as a circle with marked points, called vertices. The vertices are labeled by $V$ (preserving the cyclic ordering). The arcs between consecutive vertices are called {\em edges}. There are $k$ wavy lines in the interior of the diagram depicting the propagators, whose endpoints lie on the edges: specifically, a propagator $p =(i,j)$ has one endpoint on the edge joining vertices $i$ and $i+1$ and another endpoint on the edge joining $j$ and $j+1$, where $i+1$ and $j+1$ denote the successor of $i$ and $j$ respectively in the cyclic order on $V$.

@@

Note also that if two propagators have the same supporting edges, or if a propagator is supported on adjacent edges then the Wilson loop diagram is not admissible.

Further if we take any admissible Wilson loop diagram and remove the non-supporting vertices, we obtain a weakly admissible Wilson loop diagram that may or may not be admissible itself.

@@

Note that by Definition \ref{admisdfn}, any single propagator and its support defines an exact subdiagram. These are called \emph{trivial} subdiagrams.

@@

\vspace{3pt}
\begin{eg}
The Wilson loop diagram $W$ in Example~\ref{eg of WLD} is an example of an admissible diagram. It has a single non-supporting vertex, namely the vertex labeled 7, so we have $F(\emptyset) = \{7\}$. For $p = (5, 8)$, we have $V(p) = \{5,6,8,1\}$ and $F(p) = \{6, 7,8\}$: the vertices 1 and 5 are excluded from $F(p)$ since they support other propagators as well, and 7 is included (despite not being in the support of $p$) since it supports no propagators at all. Finally, the subdiagram $(\{(2, 4), (1, 4)\}, \{1, 2, 3, 4, 5, \})$ is an example of an exact subdiagram; in fact it is the only non-trivial exact subdiagram in $W$.
\end{eg}

@@

\begin{dfn}\label{equivdfn} 
Let $W = (\cP,[n])$ and $W' = (\cP',[n])$ be two admissible subdiagrams. We say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with the same number of propagators and suppored on the same vertices; $|P| = |P'|$ and  $V(P) =  V(P')$.
\item The complementary subdiagrams (i.e. all other propagators and vertices) are identical, i.e. $(\cP \setminus P, [n]) = (\cP' \setminus P', [n])$.
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

\begin{eg} \label{eg:equivdiags}
Note that we genuinely do need to take the transitive closure of this relation in order to obtain an equivalence relation. For example, consider the following three Wilson loop diagrams: 
\bas
W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $q$}
	\drawlabeledprop{2}{0}{4}{-1}{\footnotesize  $p$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize \quad $r$}
    \drawlabeledprop{5}{1}{7}{0}{\footnotesize \quad $s$}
		\end{scope}
	\end{tikzpicture} \quad
W_2 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $q$}
	\drawlabeledprop{2}{0}{4}{-1}{\footnotesize $p$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize \quad $u$}
    \drawlabeledprop{6}{0}{8}{-1}{\footnotesize $v$}
		\end{scope}
	\end{tikzpicture} \quad
W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $a$}
	\drawlabeledprop{1}{1}{3}{0}{\footnotesize \quad $b$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize \quad $u$}
    \drawlabeledprop{6}{0}{8}{-1}{\footnotesize $v$}
		\end{scope}
	\end{tikzpicture} .
\eas

We have $W_1 \sim W_2$ because they are identical except on the exact subdiagrams 
\[(\{r, s\}, \{5,6,7,8,1\}) \text{ and }(\{u, v\}, \{5,6,7,8,1\})\] respectively. We also have $W_2 \sim W_3$ due to the exact subdiagrams 
\[(\{p, q\}, \{1,2,3,4,5\}) \text{ and }(\{a, b\}, \{1,2,3,4,5\}).\] 
However, the difference between $W_1$ and $W_3$ is the union of two exact subdiagrams, which is not itself exact.

As described in the introduction, Wilson loop diagrams come from the study of SYM $N = 4$ in the physics literature and in particular, see \cite{Adamo:2012xe} for details, this gives that each diagram $W$ is associated to a matrix of data $C(W)$, which we now describe.

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$, we can consider the restriction $C(W)|_V$ of $C(W)$ consisting of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, any subset $V$ that has fewer than $|V|$ rows of non-zero entries does not have full rank. This is equivalent to saying that if $V$ is a set of columns of $C(W)$ with $|\Prop(V)| < |V|$, $V$ is a dependent set. Any set subset of columns containing such a $V$ cannot be independent. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$.

@@

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ (from the definition of $F(P)$) and no subset of $S$ supports fewer propagators than the number of vertices it contains (by Theorem~\ref{thm WLD defines matroid}). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup \{v\}$ also satisfies this independence condition. Furthermore, since $S \cup \{v\}$ is independent, by the first point in this lemma, $|S| = \rk(S) = \rk F(P)$. Similarly $|S \cup \{v\}| = |S|+1 = \rk (S \cup \{v\}) \leq \rk(F(P) \cup \{v\})$.  Thus $\rk(F(P) \cup\{v\}) = \rk(F(P)) + 1$, as required.
\end{proof}

@@

This section explores the correspondence between Wilson loop diagrams and positroids, established in \cite{wilsonloop}. In that paper, the authors show that admissible Wilson loop diagrams define matroids that are also positroids, Theorem \ref{put it back in then ref}. Furthermore, they show that if two Wilson loop diagrams are equivalent, they define the same matroid [ref thm].

In this section, we take this analysis further, showing the converse of Theroem [ref] from loc. cit. (see Theorem \ref{same matroid iff equiv}. We also enumerate the number of diagrams that map to an particular matroid, Corollary \ref{number of equiv diagrams}. To do this, we establish a connectection between Wilson loop diagrams and polygon dissections, subsection \ref{sec: polygon partitions}, and prove some matroidal properties of exact subdiagrams in subsection \ref{sec: exact diagram matroidal props}  including that they correspond to uniform matroids (Theorem \ref{exactuniformthm}), and the associated matroid can be written as as a contraction by the complementary propagator flat (Theorem \ref{exact diagrams contractions}).\sanote{need to discuss positroids, and whether positroids should appear earlier so that they can be discussed here.}

@@

\begin{proof}
Since the vertices of $\tau(W)$ are labeled by the edges of $W$, which are cyclically ordered, this gives an ordering to the vertices and the outer face of $\tau(W)$ is a cycle.  Since $W$ does not admit any propagators of the form $p = (i, i+1)$, there is exactly one edge connecting any two adjacent edges of $\tau(W)$. There do not exist two propagators $p,q$ such that both $p$ and $q$ start at edge $i$ and end at edge $j$. Therefore, no other two vertices of $\tau(W)$ can be connected by more than one edge.  Therefore $\tau(W)$ is a simple graph. Finally, the embedding of $\tau(W)$ is induced from the embedding of the graph $W$ and since $W$ is admissible, no pairs of propagators cross. Therefore, it is a planar embedding. 
 \end{proof}

@@

By \emph{splitting} a vertex $v$ we will mean replacing $v$ by new vertices $v_1, v_2,\ldots, v_{\text{deg}(v)}$ such that each $v_i$ has exactly one neighbour and the union of the of the $v_i$ is the neighbourhood\footnote{The neighbourhood of a vertex is the set of adjacent vertices} of $v$.

@@

Split every vertex of $T(W)$ which has degree $>3$ (see Figure~\ref{fig:vertex splitting in the dual graph} for an example of this). That is, split $T(W)$ at every vertex corresponding to a non-triangle face of $\tau(W)$. 
%By construction, each face of $\tau(W)$ is a polygon. Therefore, each vertex of $T(W)$ is at least 3 valent (one for each edge of the polygon). \hlfix{If split each vertex of $T(W)$ if it is strictly greater than $3$ valent.}{???} (In other words, we fail to split exactly when the face is a triangle).
We claim that the connected components of $T(W)$ correspond to the decomposition of $\tau(W)$ into maximal triangulated pieces and edges originally in the polygon of $\tau(W)$.  To see this, let $f$ be the forest thus obtained.
The vertices of $f$ either have degree 1 or 3.
%, (if they correspond leaf vertices) or 3,  (if they correspond to triangular faces).
Trees of $f$ with no trivalent vertices correspond to either edges in the polygon of $\tau(W)$, if they were originally leaves of $T(W)$, or to maximal trivial triangulated pieces.
Splitting at all the faces that are not triangles ensures maximality of the decomposition. If the splitting were not maximal, then one could add a triangle to a connected component of the splitting, but this would imply that that splitting happened at a degree $3$ vertex.

@@

To see uniqueness, consider a different maximal decomposition of $\tau(W)$. This induces a splitting on $T(W)$, where each connected component of the new decomposition corresponds to a subtree. Call this forest $f'$. Since $f' \neq f$, there are two trees, $t$ and $t'$ in $f$ and $f'$ respectively that are distinct, but share at least one edge of $T(W)$. Since $f'$ is also maximal, $t'$ is not a subtree of $t$. Therefore, the edges of $t'$ can be found in at least two trees in the forest $f$. In particular, there is a non-leaf vertex $v$ in $t'$ that corresponds to a split vertex of $T(W)$ in the original decomposition.
This implies that $v$ has degree greater than $3$ in $T(W)$, and thus the corresponding face of $\tau(W)$ is not a triangle. In other words, the piece of $\tau(W)$ corresponding to $f'$ is not a triangulation. 
\end{proof}

@@

Next, we introduce the idea of retriangulations. 
\begin{dfn} \label{retriangulation}
  If $W = (\cP, [n])$ and $W' = (\cP', [n])$ are weakly admissible Wilson loop diagrams, then $\tau(W)$ and $\tau(W')$  differ by a \emph{retriangulation} if there is a bijection $f$ between the maximal triangulated pieces of $\tau(W)$ and the maximal triangulated pieces of $\tau(W')$ such that for any maximal triangulated piece $t$ of $\tau(W)$ the vertex sets of $t$ and $f(t)$ are the same.
\end{dfn}
More casually this definition says that the maximal triangulated pieces of $\tau(W)$ and $\tau(W')$ are the same (so far as their vertex sets are concerned) but that some or all of the triangulated pieces may be differently triangulated while the region being triangulated by each piece (which is defined by the vertex set) remains unchanged. This defintion of retriangulation makes sense for any two polygon dissections of the same polygon, not just for those in the image of $\tau$.

@@

For an admissible Wilson loop diagram $W$, every triangulated piece, $t$, of $\tau(W)$ corresponds to a subdiagram of $W$ by taking the set of propagators $P$ corresponding to edges of $t$ and then taking the subdiagram $W_P = (P, V(P))$.  Conversely, given a subdiagram $W_P = (P, V(P))$ of $W$ we can obtain a subgraph of $\tau(W)$, called $t$, as follows:
\begin{itemize}
\item The vertex set of the subgraph $t$ is
  \begin{itemize}
  \item the vertices of $\tau(W)$ corresponding to edges of $W$ whose end points are elements of $V(P)$ which are cyclically consecutive in $[n]$. (This is significantly easier to interpret in picture form: see Example~\ref{eg:subtrees} below.)
  \end{itemize}
\item The edge set of the subgraph is
  \begin{itemize}
  \item the edges of $\tau(W)$ corresponding to propagators of $P$,
  \item the outer edges of $\tau(W)$ for which both their end points are in the vertex set.
  \end{itemize}
\end{itemize}

\begin{eg}\label{eg:subtrees}
Consider the following Wilson loop diagram $W$ and its corresponding polygon dissection $\tau(W)$. 
\bas W = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{10}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
	\drawprop{4}{1}{8}{0}
	\drawprop{8}{1}{1}{-1}
	\drawprop{5}{0}{7}{0}
%\draw (0,-2) --(0,2); horiontal
%\draw (-2,0) --(2,0);
%\draw (0,-2) node {\footnotesize $-y$};
%\draw (-2,0) node {\footnotesize $-x$};
\draw (-.3,1.) node {\footnotesize $p$};
\draw (-.2,.65) node {\footnotesize $q$};
\draw (0,-.4) node {\footnotesize $r$};
\draw (.7, 0) node {\footnotesize $s$};
		\end{scope}
	\end{tikzpicture}  \; \quad
\tau(W) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{10}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{2}{4}
    \drawchord{4}{8}
    \drawchord{8}{1}
    \drawchord{5}{7}
\draw (-.3,1.) node {\footnotesize $p$};
\draw (-.2,.65) node {\footnotesize $q$};
\draw (0,-.4) node {\footnotesize $r$};
\draw (.7, 0) node {\footnotesize $s$};
	\end{scope}
	\end{tikzpicture} \eas

>>>>

In \cite{wilsonloop} Agarwala and Amat showed that every admissible Wilson loop diagrams can be identified with a unique positroid cell, and observed that this mapping is neither one-to-one nor onto. In this paper we give a simple characterization of this lack of injectivity in terms of an equivalence relation on the Wilson loop diagrams, and enumerate the number of different Wilson loop diagrams that correspond to each positroid cell. We also study the lack of surjectivity, and show that the number of distinct positroid cells associated to Wilson loop diagrams of a given size is given by counting the directions in an associahedron.

The second paper \cite{generalcombinatoricsII} in this series identifies precisely which positroid cell is associated to a given Wilson loop diagram, by giving an algorithm for passing from a Wilson loop diagram to the Grassmann necklace of the associated positroid cell. In this manner we give a direct path from Wilson loop diagram to positroid cell, circumventing the painstaking process of examining sets of bases. With this identification in hand, we also prove that any Wilson loop diagram with $k$ propagators corresponds to a $3k$ dimensional positroid cell, and characterize the volume forms that Wilson loop diagrams associate to each positroid cell.

@@

Meanwhile, there is a separate body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. As noted above, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of the positive Grassmannian $\Gr(k,n)$ \cite{wilsonloop}. In (separate) papers with Fryer and Marcott, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmannians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

However, the two approaches are not identical: for example, in the case of Wilson loop diagrams with $k=2$ and $n=6$, past work of Agarwala and Fryer shows that the manifold is not contractible \cite{casestudy}. More generally, it has been shown that the spaces defined by the Amplituhedron and by Wilson loop diagrams are not the same \cite{Amplituhedronsquared}. In fact, it is in some ways more natural to consider Deodhar cells of $\mathbb{G}(k, n+1)$ for Wilson loop diagrams; in this case, one gets that the manifold represented by the diagrams is not orientable \cite{non-orientable}. \note{Check with Cam.}

@@

Wherever possible, we drop the subscripts on propagators and their identifying vertices. Propagators are undirected, so $p = (i,j) = (j,i)$. As in previous work, we follow the convention of writing $p = (i,j)$ with  $i +1 \leq j$ relative to the first vertex.

A Wilson loop diagram can also be represented visually, which provides useful intuition for many of the results that follow. We depict the diagram $(\cP,V)$ as a circle with marked points, called {\em vertices}; these vertices are labeled by $V$ (preserving the cyclic ordering). The arcs between consecutive vertices are called {\em edges}. There are also $k$ wavy lines in the interior of the diagram depicting the propagators, whose endpoints lie on the edges: specifically, a propagator $p =(i,j)$ has one endpoint on the edge joining vertices $i$ and $i+1$ and another endpoint on the edge joining $j$ and $j+1$, where $i+1$ and $j+1$ denote the successor of $i$ and $j$ respectively in the cyclic order on $V$.

@@

Note also that if two propagators have the same supporting edges, or if a propagator is supported on adjacent edges, then the Wilson loop diagram is not admissible.

Further, if we take any admissible Wilson loop diagram and remove the non-supporting vertices, we obtain a weakly admissible Wilson loop diagram that may or may not be admissible itself.

@@

Note that by Definition \ref{admisdfn}, any single propagator and its support defines an exact subdiagram. These are called \emph{trivial} exact subdiagrams.

@@

\vspace{3pt}
\begin{eg}
The Wilson loop diagram $W$ in Example~\ref{eg of WLD} is an example of an admissible diagram. It has a single non-supporting vertex, namely the vertex labeled 7, so we have $F(\emptyset) = \{7\}$. For $p = (5, 8)$, we have $V(p) = \{5,6,8,1\}$ and $F(p) = \{6, 7,8\}$: the vertices 1 and 5 are excluded from $F(p)$ since they support other propagators as well, and 7 is included (despite not being in the support of $p$) since it supports no propagators at all. Finally, the subdiagram $(\{(2, 4), (1, 4)\}, \{1, 2, 3, 4, 5\})$ is an example of an exact subdiagram; in fact it is the only non-trivial exact subdiagram in $W$.
\end{eg}

@@

\begin{dfn}\label{equivdfn} 
Let $W = (\cP,[n])$ and $W' = (\cP',[n])$ be two admissible subdiagrams. We say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with the same number of propagators and supported on the same vertices, i.e. $|P| = |P'|$ and  $V(P) =  V(P')$.
\item The complementary subdiagrams (i.e. all other propagators and vertices) are identical, i.e. $(\cP \setminus P, [n]) = (\cP' \setminus P', [n])$.
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

\begin{eg} \label{eg:equivdiags}
Note that we genuinely do need to take the transitive closure of this relation in order to obtain an equivalence relation. For example, consider the following three Wilson loop diagrams: 
\bas
W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $q$ \quad}
	\drawlabeledprop{2}{0}{4}{-1}{\footnotesize  $p$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize $r$ \ \ }
    \drawlabeledprop{5}{1}{7}{0}{\footnotesize \quad \ $s$}
		\end{scope}
	\end{tikzpicture} \quad
W_2 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $q$ \quad}
	\drawlabeledprop{2}{0}{4}{-1}{\footnotesize $p$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize $u$ \ \ }
    \drawlabeledprop{6}{0}{8}{-1}{\footnotesize $v$ \ }
		\end{scope}
	\end{tikzpicture} \quad
W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $a$ \quad}
	\drawlabeledprop{1}{1}{3}{0}{\footnotesize  $b$ \ \ \ }
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize $u$ \ \ }
    \drawlabeledprop{6}{0}{8}{-1}{\footnotesize $v$ \quad }
		\end{scope}
	\end{tikzpicture} .
\eas

$W_1$ and $W_2$ are related because they are identical except on the exact subdiagrams 
\[(\{r, s\}, \{5,6,7,8,1\}) \text{ and }(\{u, v\}, \{5,6,7,8,1\})\] respectively. Similarly, $W_2$ and $W_3$ are related due to the exact subdiagrams 
\[(\{p, q\}, \{1,2,3,4,5\}) \text{ and }(\{a, b\}, \{1,2,3,4,5\}).\] 
However, the difference between $W_1$ and $W_3$ is the union of two exact subdiagrams, which is not itself exact.

\end{eg}
\medskip

As described in the introduction, Wilson loop diagrams come from the study of SYM $N = 4$ in the physics literature. In particular (see \cite{Adamo:2012xe} for details) this gives that each diagram $W$ is associated to a matrix of data $C(W)$, which we now describe.

@@

Throughout, we take the {\em matroid defined by $W$} to be the matroid $M(W)$ of Theorem \ref{thm WLD defines matroid}. Note from Definition~\ref{def:CWmatrix} that the vertices of the diagram $W$ correspond to columns of the associated matrix $C(W)$. Therefore, for any subset of vertices $V \subseteq [n]$, we can consider the restriction $C(W)|_V$ of $C(W)$ consisting of the columns corresponding to $V$. Since the non-zero entries of $C(W)$ are independent real variables, if $C(W)|_V$ has fewer than $|V|$ rows of non-zero entries then it not have full rank. This is equivalent to saying that if $V$ is a set of columns of $C(W)$ with $|\Prop(V)| < |V|$, then $V$ is a dependent set. Any set of columns containing such a $V$ cannot be independent. Comparing this to Theorem \ref{thm:WLDmatroid}, we see that $M(W)$ is also the matroid realized by $C(W)$.

@@

To prove (4), we need to show that $F(P)$ is maximally dependent. If $F(P) = [n]$ then this is automatic, so suppose not and let $v \in [n] \setminus F(P)$. In other words, $v \in V(P^c)$ and so $v$ supports some propagator $q \not\in P$.  Let $S \subseteq F(P)$ be an independent set of maximal size. Then $\Prop(S) \subseteq P$ (from the definition of $F(P)$) and no subset of $S$ supports fewer propagators than the number of vertices it contains (by Theorem~\ref{thm WLD defines matroid}). Since $v$ supports a new propagator $q \not\in P$, the set $S \cup \{v\} \subseteq F(P) \cup \{v\}$ also satisfies this independence condition.

Since $S$ is independent, it follows from the first point in this lemma that $|S| = \rk(S) = \rk F(P)$. Similarly, 
\[|S \cup \{v\}| = |S|+1 = \rk (S \cup \{v\}) \leq \rk(F(P) \cup \{v\}) \leq \rk F(P) + 1.\]
Thus $\rk(F(P) \cup\{v\}) = \rk F(P) + 1$, as required.
\end{proof}

@@

This section explores the correspondence between Wilson loop diagrams and positroids that was established in \cite{wilsonloop}. In that paper, the authors show that admissible Wilson loop diagrams define matroids that are also positroids, Theorem \ref{put it back in then ref}. Furthermore, they show that if two Wilson loop diagrams are equivalent, they define the same matroid [ref thm].

In this section we develop this analysis further, showing the converse of Theorem [ref] from loc. cit. (see Theorem \ref{same matroid iff equiv}). We also enumerate the number of diagrams that map to an particular matroid in Corollary \ref{number of equiv diagrams}. To do this, we establish a connection between Wilson loop diagrams and polygon dissections in subsection \ref{sec: polygon partitions}, and uncover some matroidal properties of exact subdiagrams in subsection \ref{sec: exact diagram matroidal props}. This includes the fact that exact subdiagrams correspond to uniform matroids (Theorem \ref{exactuniformthm}), and that the associated matroid of an exact subdiagram can be written as as a contraction by the complementary propagator flat (Theorem \ref{exact diagrams contractions}).\sanote{need to discuss positroids, and whether positroids should appear earlier so that they can be discussed here.}

@@

\begin{proof}
Since the vertices of $\tau(W)$ are labeled by the edges of $W$, which are cyclically ordered, this gives an ordering to the vertices and the outer face of $\tau(W)$ is a cycle.  Since $W$ does not admit any propagators of the form $p = (i, i+1)$ due to its admissibility, there is exactly one edge connecting any two adjacent edges of $\tau(W)$. Similarly, there does not exist two propagators $p,q$ such that both $p$ and $q$ start at edge $i$ and end at edge $j$. Therefore, no other two vertices of $\tau(W)$ can be connected by more than one edge.  Therefore $\tau(W)$ is a simple graph. Finally, the embedding of $\tau(W)$ is induced from the embedding of the graph $W$ and since $W$ is admissible, no pairs of propagators cross. Therefore, it is a planar embedding. 
 \end{proof}

@@

By \emph{splitting} a vertex $v$ we will mean replacing $v$ by new vertices $v_1, v_2,\ldots, v_{\text{deg}(v)}$ such that each $v_i$ has exactly one neighbour and the union of the $v_i$ is the neighbourhood\footnote{The neighbourhood of a vertex is the set of adjacent vertices} of $v$.

@@

Now split every vertex of $T(W)$ which has degree $>3$; see Figure~\ref{fig:vertex splitting in the dual graph} for an example of this. That is, we split $T(W)$ at every vertex corresponding to a non-triangle face of $\tau(W)$. 
%By construction, each face of $\tau(W)$ is a polygon. Therefore, each vertex of $T(W)$ is at least 3 valent (one for each edge of the polygon). \hlfix{If split each vertex of $T(W)$ if it is strictly greater than $3$ valent.}{???} (In other words, we fail to split exactly when the face is a triangle).

We claim that the connected components of $T(W)$ correspond to the decomposition of $\tau(W)$ into maximal triangulated pieces and edges originally in the polygon of $\tau(W)$.  To see this, let $f$ be the forest thus obtained.
The vertices of $f$ either have degree 1 or 3.
%, (if they correspond leaf vertices) or 3,  (if they correspond to triangular faces).
Trees of $f$ with no trivalent vertices correspond to either edges in the polygon of $\tau(W)$ (if they were originally leaves of $T(W)$) or to maximal trivial triangulated pieces.
Splitting at all the faces that are not triangles ensures maximality of the decomposition. If the splitting were not maximal, then one could add a triangle to a connected component of the splitting, but this would imply that that splitting happened at a degree $3$ vertex.

@@

To see uniqueness, consider a different maximal decomposition of $\tau(W)$. This induces a splitting on $T(W)$, where each connected component of the new decomposition corresponds to a subtree. Call this forest $f'$. Since $f' \neq f$, there are two trees $t$ and $t'$ in $f$ and $f'$ respectively that are distinct, but share at least one edge of $T(W)$. Since $f'$ is also maximal, $t'$ is not a subtree of $t$. Therefore, the edges of $t'$ can be found in at least two trees in the forest $f$. In particular, there is a non-leaf vertex $v$ in $t'$ that corresponds to a split vertex of $T(W)$ in the original decomposition.
This implies that $v$ has degree greater than $3$ in $T(W)$, and thus the corresponding face of $\tau(W)$ is not a triangle. In other words, the piece of $\tau(W)$ corresponding to $f'$ is not a triangulation. 
\end{proof}

@@

Next, we introduce the idea of retriangulations. 
\begin{dfn} \label{retriangulation}
  If $W = (\cP, [n])$ and $W' = (\cP', [n])$ are weakly admissible Wilson loop diagrams, then $\tau(W)$ and $\tau(W')$  differ by a \emph{retriangulation} if there is a bijection $f$ between the maximal triangulated pieces of $\tau(W)$ and the maximal triangulated pieces of $\tau(W')$ such that for any maximal triangulated piece $t$ of $\tau(W)$ the vertex sets of $t$ and $f(t)$ are the same.
\end{dfn}
More casually this definition says that the maximal triangulated pieces of $\tau(W)$ and $\tau(W')$ are the same (so far as their vertex sets are concerned) but that some or all of the triangulated pieces may be differently triangulated while the region being triangulated by each piece (which is defined by the vertex set) remains unchanged. This definition of retriangulation makes sense for any two polygon dissections of the same polygon, not just for those in the image of $\tau$.

@@

For an admissible Wilson loop diagram $W$, every triangulated piece $t$ of $\tau(W)$ corresponds to a subdiagram of $W$ obtained by taking the set of propagators $P$ corresponding to edges of $t$ and then taking the subdiagram $W_P = (P, V(P))$.  Conversely, given a subdiagram $W_P = (P, V(P))$ of $W$ we can obtain a subgraph of $\tau(W)$, called $t$, as follows:
\begin{itemize}
\item The vertex set of the subgraph $t$ is
  \begin{itemize}
  \item the vertices of $\tau(W)$ corresponding to edges of $W$ whose end points are elements of $V(P)$ which are cyclically consecutive in $[n]$. (This is significantly easier to interpret in picture form: see Example~\ref{eg:subtrees} below.)
  \end{itemize}
\item The edge set of the subgraph is
  \begin{itemize}
  \item the edges of $\tau(W)$ corresponding to propagators of $P$,
  \item the outer edges of $\tau(W)$ for which both their end points are in the vertex set.
  \end{itemize}
\end{itemize}

\begin{eg}\label{eg:subtrees}
Consider the following Wilson loop diagram $W$ and its corresponding polygon dissection $\tau(W)$. 
\bas W = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{10}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
	\drawprop{4}{1}{8}{0}
	\drawprop{8}{1}{1}{-1}
	\drawprop{5}{0}{7}{0}
%\draw (0,-2) --(0,2); horiontal
%\draw (-2,0) --(2,0);
%\draw (0,-2) node {\footnotesize $-y$};
%\draw (-2,0) node {\footnotesize $-x$};
\draw (2.7*\angle:0.8*\radius) node {\footnotesize $p$};
\draw (3.3*\angle:0.45*\radius) node {\footnotesize $q$};
\draw (0,-.4) node {\footnotesize $r$};
\draw (.7, 0) node {\footnotesize $s$};
		\end{scope}
	\end{tikzpicture}  \; \qquad
\tau(W) = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawpolypart{10}{1.5}
    \drawnumbersshift
    \drawchord{1}{4}
    \drawchord{2}{4}
    \drawchord{4}{8}
    \drawchord{8}{1}
    \drawchord{5}{7}
\draw (2.7*\angle:0.8*\radius) node {\footnotesize $p$};
\draw (3.3*\angle:0.47*\radius) node {\footnotesize $q$};
\draw (0,-.4) node {\footnotesize $r$};
\draw (.7, 0) node {\footnotesize $s$};
	\end{scope}
	\end{tikzpicture} \eas
